---
title: "Getting and Cleaning Data"
author: "Ehsan Ullah"
date: "10/9/2019"
output: html_document
---

## Week 1
### Video Lecture 9

### The data.table Package

```{r Create data tables like data frames}
library(data.table)

DF = data.frame(x=rnorm(9), y=rep(c("a","b","c"), each=3), z=rnorm(9))

head(DF, 3)

DT = data.table(x=rnorm(9), y=rep(c("a","b","c"), each=3), z=rnorm(9))

head(DT, 3)

tables() #shows the details about all the data tables in the workspace. different than table() command. it has an S in the end.

```

### Subsetting rows

```{r subsetting rows}
DT[2,]

DT[DT$y=="a",]

DT[c(2,3)] # single index will extract only rows and not columns.

```

### Subsetting columns
```{r subsetting columns}
DT[,c(2,3)] #in course it says it does something different but infact it really extracted second and third columns

```

### Column subsetting in data.table()
- Subsetting function in R is modified for data.table().
- The argument you pass after "comma" is called an "expression".
- In R an expression is collection of statments enclosed in curly brackets.

```{r}
{ 
        x = 1
        y = 2
}

k = {print(10); 5}

print(k)
```

### Calculating values for variables with expressions.
```{r}
DT[, list(mean(x), sum(z))] # No quotation marks required.

DT[,table(y)]

```

### Adding new columns
```{r}

DT[, w:=z^2]

head(DT, 3)

#Warning: This will change the original dataset DT.

DT[, y:=2]

head(DT, 3)
#To create a copy of a data.table, use copy function.
DT4 <- copy(DT) 

```

```{r}


```

```{r}


```

### Multiple operations
```{r}

DT[, m := {tmp <- (x + z); log2(tmp+5)}]

head(DT, 4)

```
### plyr like operations
```{r}
DT[, a:= x>0]

head (DT, 3)

# We can apply different functions after grouping the variables by a. for example

DT[, b:=mean(x+w), by=a]

head(DT) # We can see two different types of means, one for TRUE and one for FALSE of a. 

```


### Special variables
```{r}
# .N An integer, length 1, contaning the nubmers
DT3 <- data.table(x=sample(letters[1:3], 1e5, TRUE))

DT3[, .N, by=x]

```

### Keys
```{r}
DT5 <- data.table(x=rep(c("a","b","c"), each=1000), y=rnorm(900))

head(DT5, 5)

setkey(DT5, x)

DT5['c'] # it will look ONLY for c in x, which is the key).
```

### Joins
```{r}
DT6 <- data.table(x=c("a", "b", "c", "dt1"), y=1:4)
DT7 <- data.table(x=c("a", "b", "c", "dt2"), z=5:7)

setkey(DT6, x); setkey(DT7, x)

merge(DT6, DT7)

```

### Reading time faster in data.table()

```{r}
big_df <- data.frame(x=rnorm(1E6), y=rnorm(1E6))

file <- tempfile()

write.table(big_df, file=file, row.names = FALSE, col.names = TRUE, sep= "\t", quote=FALSE)

system.time(fread(file))


system.time(read.table(file, header=TRUE, sep="\t")) # This is really slow.


```

### Remove unncecessary data.

```{r}
rm(big_df)
rm(DT3)
rm(DT5)
rm(file)
```